Scoring epidemiological forecasts on transformed scales

### 2023-06-26 reply_to_reviews.tex

\newcommand{\indented}{\setlength{\leftskip}{1cm}}
\newcommand{\notindented}{\setlength{\leftskip}{0cm}}

\begin{document}

%\black \textbf{Black: our comments}
%\red \textbf{Reviewer, not addressed}
%\blue \textbf{Reviewer, addressed}

\black
We thank the reviewers for their feedback and thoughtful comments which we feel have helped improve the manuscript substantially. Please see our responses to the individual comments as well as explanations on the changes we have made to the manuscript below. 
In addition to the changes outlined below we some general edits to the paper. We changed the title to make it more concise, we altered the order of some sections and shortened the introduction a bit, which became quite lengthy after incorporating all the suggestions.

@@

In addition, I suspect there are a few readers (such as myself!) who are more familiar with more ‘basic’ scores such as e.g. logarithmic score or arise (i.e. essentially the likelihood or related quantities). I spent a bit of time wondering about why these weren’t considered, especially as these avoid the scale issues at the centre of the present article. Such scores were only briefly considered in the discussion, in the second-to-last paragraph. I think it would be helpful to mention/emphasise up front some of these ‘classic’ scores that don’t have the same issues with scaling and then discuss why they aren’t used, e.g. due to ‘robustness’ or other issues (e.g. lack of full probability distributions from forecasts).

\black
We updated the paragraph, which now reads:

\indented

Proper scoring rules are constructed such that they encourage honest forecasting and cannot be `gamed' or `cheated'. 
Assuming that the forecaster's actual best judgement corresponds to a predictive distribution $F$, a proper score is constructed such that if $F$ was the data-generating process, no other distribution $G$ would yield a better expected score. A scoring rule is called \textit{strictly} proper if there is no other distribution that under $F$ achieves the \textit{same} expected score as $F$, meaning that any deviation from $F$ leads to a worsening of expected scores. Forecasters (anyone or anything that issues a forecast) are thus incentivised to report their true belief $F$ about the future. 
Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} and the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}. The log score is the predictive log density or probability mass evaluated at the observed value. It is supported by the likelihood principle \citep{Winkler1996} and has many desirable theoretical properties; however, the particularly severe penalties it assigns to occasional misguided forecasts make it little robust \citep{bracherEvaluatingEpidemicForecasts2021}. Moreover, it is not easily applied to forecasts reported as samples or quantiles, as used in many recent disease forecasting efforts. It is nonetheless occasionally used in epidemiology (see e.g., \citealt{heldProbabilisticForecastingInfectious2017, Johansson2019}), but in recent years the CRPS and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021} have become increasingly popular.

@@

\blue
\subsubsection{Log vs other transformations}
The authors take the position that the log transformation is generally preferable for their use case, though they discuss and compare others. Relatedly, their arguments in terms of relative error also involve an essentially log-linear (i.e. log data + additive error) form which may be pragmatic but somewhat inelegant imo.

\black
In the theoretical part of the manuscript we simplify the equations a little bit by looking at the case of a single point forecast, rather than a probabilistic forecast that comes in the form of a predictive distribution. We feel this makes the discussion more intuitive. We want to emphasise that the log-linear error is really the error that is measured by the score, not the error model that would be assumed by any forecaster making a prediction.

\blue
As the authors note, a particularly controversial issue for the log transformation is the issue of zero values. They take the standard pragmatic stance that we can use log (eps + y) instead of log y in such cases. Many communities e.g. econometrics are vehemently opposed to this, preferring e.g. quasi-Poisson regression and robust standard errors in the context of estimation. While not completely opposed to log (eps + y) myself (I have used it too!) it makes me a bit uncomfortable I still can’t help but feel that something along the lines of e.g. logarithmic/quasi-likelihood or deviance scores could be formulated that would be preferable to the log data transform. This would offer both automatic scaling and handling of zero values, in principle. However, I haven’t thought carefully enough to offer a concrete alternative and the log(eps + y) approach appears to work reasonably here. Instead, and in combination with the previous point, perhaps a bit more discussion of the potential alternatives based on quasi-likelihood-style functions rather than data transformations could be added?

\black
Thank you for raising this interesting point. The log score is indeed an alternative which is scale invariant under monotonic transformations. We mention this in the introduction and state that for the log score the question of which scale to evaluate forecasts on does not arise. In our understanding, a quasi-likelihood approach (as opposed to an actual likelihood approach as in the log score) may not actually be necessary as at least in theory full predictive distributions are to be evaluated. Developing further ideas based on some form of quasi-likelihood would likely require more in-depth investigation. In any case, developing these ideas would require more in-depth investigation and we feel this is beyond the scope of the current paper.

@@

\black
It is important to note that the equations in 2.2 do not refer to any particular model used by a forecaster or any model that was fit to data. Rather, we define the instantaneous growth rate for a single realised epidemic and then show that scoring forecasts for that on the log scale equates to scoring a forecast of the average growth rate according to this definition. The forecaster themselves who would make a forecast could assume any error model or model structure they like.

\blue
\subsection{Specific suggestions}

I realised as I was about comment on some equations that none of the equations are numbered. It would probably be good to number them :-)

\black
You are completely right, we added numbers to all equations.

@@

``Natural logarithm'', log1p, and log($a + x$) are conflated in several places, despite the discussion in Section 2.4 noting that this is an important distinction. The terminology should not be overloaded. (E.g., L20 ``Applying the log transformation''.)

\black
We changed ``Applying the log transformation'' to ``Applying a transformation of log($x + 1$)'' in L20 (in the abstract).

We slightly disagree with the notion that the terminology should not be overloaded. Not introducing another term for transformations of $\log_{e}(x + \varepsilon)$ makes the manuscript more readable and accessible to readers. In addition, our analyses suggest that adding a small offset doesn't make much of a difference in most practical applications. We nevertheless completely agree that this should be made clear to the reader and have added the following text:

\indented
For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particularly attractive transformation in the context of epidemic phenomena. We refer to this transformation as 'log-transformation' and to scores that have been computed from log-transformed forecasts and observations as scores 'on the log scale' (as opposed to scores 'on the natural scale', which involve no transformation). In the theoretical discussion in Section 2, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$. For practical applications (Section 3) we also use these terms to describe a transformation of $\log_{e}(x + a)$ with a small $a > 0$ in order to keep the terminology and notation simple. For a prediction target with strictly positive support, the CRPS after applying a log-transformation is given by

@@

\notindented

Regarding the question which transformations are permissible:

Scaling by past observations is permissible (indeed taking the log is similar to scaling by the last value), as is scaling by the population size as mentioned in the discussion. There may of course be exceptions such as dividing by zero, but otherwise this just represents a monotonic transformation, which is unproblematic in terms of propriety. Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it disturbs propriety; see Lerch et al (2017, Statistical Science, preprint: \url{https://arxiv.org/abs/1512.09244}) on the closely related topic of weighting scores with a function of the observed value. Similarly, scaling a forecast by predictions made by other forecasters may break propriety, but we are unaware of existing theoretical arguments. Scaling scores, e.g. dividing a score by the mean score for other forecasters may be acceptable in most applied settings, but we are unsure about exactly when and how this could break propriety (as is the case for example when dividing by zero).

To make this clearer in the text, we added the following paragraphs:

\indented
A natural question is which other transformations could be applied and whether resulting scores remain (strictly) proper. In principle, any transformation function can be applied simultaneously to forecasts and observations as long as the definition of the transformation is independent of the forecasts and any quantities unknown at the time of forecasting, including the observed value. This simply corresponds to a re-definition of the forecasting target. However, applying non-invertible transformations leads to a loss in information conveyed by forecasts, which we consider undesirable. The resulting score will be proper, but it may not be strictly proper anymore (as forecasts differing from the forecaster's true belief on the original scale may be identical on the transformed scale). When using the CRPS or the WIS, it seems most appropriate to use only strictly monotonic transformations such as the natural logarithm or the square root as otherwise the encoded notion of distance may become meaningless.

Some other strictly monotonic transformations that can be applied are scaling by the population size or scaling by past observations. The latter, as discussed in Section 4, is similar to applying a log-transformation, but corresponds to evaluating a forecast of multiplicative, rather than exponential growth rates. The arising issue of dividing by zero can again be solved by adding a small offset $a$. Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it can result in improper scores (see \citealt{lerchForecasterDilemmaExtreme2015} on the closely related topic of weighting scores with a function of the observed value). Similarly, scaling forecasts and observations by a function of the predictive distribution (like the predictive mean) may lead to improper scores; however, we are unaware of existing theoretical arguments on this.

When applying a transformation, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee that the score remains proper. In the case of log transforms, taking the logarithm of the CRPS values, rather than scoring the log-transformed forecasts and data, results in an improper score. We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can be seen that in the example overconfident models perform best in terms of the log WIS. We note that strictly speaking, re-scaling average scores by the average score of a baseline model or average scores across different models  to obtain skill scores likewise leads to improper scores \citep{gneitingStrictlyProperScoring2007}. The application of such skill scores, however, is established practice and considered largely unproblematic.

@@

\blue
L46 “WIS is an approximation of the CRPS” — Is it an approximation of CRPS itself or some (2x?) scaling?

\black
The weighted interval score is a weighted sum of interval scores for the individual prediction intervals. When weighing the individual interval scores for the central $1 - \alpha$ prediction intervals by 
$w = \alpha / 2$ (which is generally standard), then the WIS converges exactly to the CRPS for an increasing number of equally spaced prediction intervals. For an illustration see Figure 2 in \cite{bracherEvaluatingEpidemicForecasts2021}.

\blue
L59 “strength” — The meaning of “strength” is nonintuitive without reading the citation and understanding the ties to multiplicative interventions on R and r (the latter of which seems unrealistic to achieve). Consider surrounding with quotes, explaining more/less, relating to growth per generation, etc.

\black
We revised and shortened the introduction, so this sentence has been removed. Now, it simply reads:

Especially in their early phase, outbreaks are best conceived as exponential processes, characterized by potentially time varying reproduction numbers $R_t$ \citep{gosticPracticalConsiderationsMeasuring2020} or epidemic growth rates $r_t$ \citep{dushoffSpeedStrengthEpidemic2021}.

@@

Especially in their early phase, outbreaks are best conceived as exponential processes, characterized by potentially time varying reproduction numbers $R_t$ \citep{gosticPracticalConsiderationsMeasuring2020} or epidemic growth rates $r_t$ \citep{dushoffSpeedStrengthEpidemic2021}.

\notindented

\blue
L60: “immunity” $\xrightarrow{}$ “population immunity”?

\black
We agree with you, but have simplified the introduction such that this sentence does no longer appear.

@@

\blue
L69: This might benefit from a short justification (e.g.,
$\left| \frac{x_t e^{(r_t + \epsilon) \Delta t} - x_t e^{r_t \Delta t}}{x_t e^{(r_t - \epsilon) \Delta t} - x_t e^{r_t \Delta t}} \right| = e^{\epsilon\Delta t}$)

\black
This is a great suggestion, thank you. We updated the corresponding paragraph, which now reads:

\indented
If the true modelling task revolves around estimating and forecasting these quantities, then evaluating forecasts based on the absolute distance between forecasted and observed incidence values penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount. For illustration, consider an incidence forecast issued at time 0 and referring to time $t$ that misses the correct average growth rate $\bar{r}_t$ by either $-\epsilon$ or $+\epsilon$. Then the ratio of the resulting absolute errors on the scale of observed incidences $y_{t}$ is

\begin{equation}
\frac{\left|y_0 \exp[(\bar{r}_t - \epsilon) \times t] - y_0 \exp(\bar{r}_t t)\right|}{\left| y_0 \exp[(\bar{r}_t + \epsilon) \times t] - y_0 \exp(\bar{r}_t t) \right|} = \exp(-\epsilon t) < 1.
\end{equation}

If one is to measure the ability to forecast the underlying infection dynamics, it may thus be more desirable to evaluate errors on the scale of the growth rate directly.

@@

\blue
L272: “with the standard deviation” — Also note here that this is for approximately normal data distributions?

\black 
We now remind the reader of this assumption by writing

\indented

As (under the simplifying assumption of normality, see Section 2) the WIS/CRPS of an ideal forecaster scales with the standard deviation, a value of $\beta = 1$ would imply a quadratic median-variance relationship;

@@

L345: “divide each forecast by the forecast” — Caveat about current forecast formats should be included here / in following sentences, without (one mention of) the idea about directly soliciting forecasts of week-on-week ratios in between.

\black
We updated the relevant paragraph which now reads:

\indented

We suggested using the natural logarithm as a variance-stabilising transformation (VST). This is appropriate for variables that are approximately normally distributed and have a quadratic mean-variance relationship with $\sigma^2 = c \times \mu^2$ (this is e.g. approximately true for the negative binomoial distribution and large $\mu$). Alternatively, the square-root transformation can be appropriate in the case of a Poisson distributed variable \citep{dunnGeneralizedLinearModels2018}. Other VST like the Box-Cox \citep{boxAnalysisTransformations1964} are conceivable as well.
If one is interested in multiplicative, rather than exponential growth rates, one could, instead of applying a log transformation, convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last value that was observed at the time the forecast was made. Forecasters would then implicitly predict a separate multiplicative growth rate from today to horizon 1, 2, etc. 
Instead of dividing by the last observed value, another promising transformation would be to divide each forecast by the forecast of the previous week (and analogously for observations), in order to obtain forecasts for week-to-week growth rates. Alternatively, one could also take first differences of values on the log scale. This approach would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluating the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}. Dividing values by the previous value, unfortunately, is not feasible under the current quantile-based format of the Forecast Hubs, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate. However, it may be an interesting approach if predictive samples are available or if quantiles for weekwise growth rates have been collected. Potentially, the variance stabilising time-series forecasting literature may be a useful source of other transformations for various forecast settings.

@@

\indented
For a given solid line, the dashed line of the same colour marks the x-value that  is equal to 5 times the corresponding offset. It can be seen that for $a$ values smaller than one fifth of the transformed quantity, the effect of adding an offset is generally small. When choosing a suitable $a$, the trade-off is between staying close to the interpretation of a pure log-transformation (choosing a small $a$) and not giving excessive weights to small observations (by choosing a larger $a$, see Figure 3).

\notindented

\blue
Figure SI.7:
\begin{itemize}
    \item Also specify sorting variable\&direction on x axis. (Would a point plot vs. the sorting variable be more informative?)
    \item Very informative figure, promote to main text? Also would like to see for different “parts” of an epidemic (e.g., steady low periods, increases, peaks/plateaus, decreases). Deaths figures suggest that balancing scores between locations can be tricky, and both cases and deaths suggests that the investigated transformations can’t do this exactly.
\end{itemize}

\black
We added a statement to the legend "Locations are sorted according to the mean observed value in that location." We also followed your suggestion to make this a main figure and added some explanations in the text.

@@

\blue
1. Line 96, typo: "log-transformtation".

\black
We corrected the typo

\blue
2. Figure 3: out of curiosity, do the rankings of forecasts A and B first switch at the same observed value (somewhere around 7?) in both panels? Zooming in, it looks as though maybe the switch occurs very slightly earlier for the log-transformed scoring.

\black
You're completely right, on the natural scale, the switch happens between 6.5 and 6.6 and on the log scale between 6.2 and 6.3.

@@

\blue
4. Line 281: missing closing parenthesis, "(or $\beta_{\sqrt{}} = 0$, respectively".

\black
We corrected the typo

\blue
5. Figure 7: I can appreciate that the correlation between the natural and logarithmic scores decreases over the forecast horizon, as the absolute forecast error grows. But I found panel B more challenging to interpret. I gather the key message is that model rankings are quite consistent between the natural and log scales (panel A) even though the absolute scores differ markedly between the two scales — an effect that increases over the forecast horizon (panel B). Perhaps adding a sentence to this effect in the accompanying text (lines 292-295) would help other readers to avoid my confusion.

\black
You'r right, we adapted the text and hopefully made it more clear:

For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure 8A). While rankings between forecasters remain similar for a single forecast, this is not true anymore when looking at rankings obtained after averaging scores across multiple forecasts made at different times or in different locations. As discussed earlier, scores on the natural and on the log scale penalise errors very differently, e.g. when looking at performance during peaks or troughs. When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models therefore change considerably (Figure 8B).

@@

\black
While more research is needed to conclude this with more confidence, we generally agree with the statement. We would also expect ensemble forecasts to do well (as they have done in the post) regardless of whether they are evaluated on the log or on the natural scale.

\blue
8. Lines 327-329: "A potential downside is that forecast evaluation is unreliable in situations where observed values are zero or very small."

I feel there's an argument to be made here that rather than being a potential limitations, it could be considered an accurate reflection of the inherent uncertainty about the future course of an epidemic when case numbers are very small. This transformation also provides a valuable benefit in these situations, as the authors note a few sentences later: "It also gives higher weight to another type of situation one may care about, namely one in which numbers start to rise from a previously low level". This neatly illustrates the importance (as highlighted by the authors in this manuscript) of using multiple scoring rules to evaluate and compare forecasts.

\black
We think this is a valid point and added a sentence to that effect:

### 2023-06-26 main.tex
fix small latex issue

\begin{table}
\centering
\begin{tabular}{ccrrrrrr}
\toprule
Horizon & Target & $\alpha$ & $\beta$ & $\alpha_\sqrt{\ }$ & $\beta_\sqrt{\ }$ & $\alpha_\log$ & $\beta_\log$ \\
\midrule
% all & all & -1.093 & 0.963 & -0.352 & 0.201 & 0.391 & 0.001\\
% \addlinespace
% all & Cases & 0.036 & 0.858 & 0.043 & 0.201 & 0.751 & -0.033\\
% all & Deaths & -0.884 & 0.868 & 0.273 & 0.121 & 0.436 & -0.023\\
% \addlinespace
% 1 & all & -1.402 & 0.923 & 0.320 & 0.088 & 0.318 & -0.014\\
% 2 & all & -1.221 & 0.967 & 0.112 & 0.164 & 0.364 & -0.003\\
% 3 & all & -1.001 & 0.984 & -0.094 & 0.241 & 0.410 & 0.008\\
% 4 & all & -0.757 & 0.986 & 0.000 & 0.299 & 0.469 & 0.015\\
1 & Cases & -0.862 & 0.876 & 0.790 & 0.087 & 0.433 & -0.024\\
2 & Cases & -0.243 & 0.877 & 0.959 & 0.162 & 0.660 & -0.031\\
3 & Cases & 0.372 & 0.855 & 1.109 & 0.238 & 0.882 & -0.037\\
4 & Cases & 0.816 & 0.837 & 1.645 & 0.296 & 1.009 & -0.036\\
\addlinespace
1 & Deaths & -1.146 & 0.832 & 0.457 & 0.048 & 0.376 & -0.035\\
2 & Deaths & -0.981 & 0.867 & 0.443 & 0.084 & 0.416 & -0.028\\
3 & Deaths & -0.807 & 0.885 & 0.349 & 0.131 & 0.453 & -0.019\\
4 & Deaths & -0.602 & 0.891 & 0.125 & 0.194 & 0.501 & -0.011\\
\bottomrule
\end{tabular}
\caption{Coefficients of three regressions for the effect of the magnitude of the median forecast on expected scores. The first regression was 
$\log[\text{WIS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], $ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
$\text{WIS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},$ where $F_{\log}$ is the predictive distribution for $\log y$. The third one was $\text{WIS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{sqrt{\ }} + \beta_{sqrt{\ }} \cdot \sqrt{(\text{median}(F))},$ where $F_{\sqrt{\ }}$ is the predictive distribution for $\sqrt{y}$.
}
\label{tab:HUB-regression}
\end{table}

>>>>

\begin{table}
\centering
\begin{tabular}{ccrrrrrr}
\toprule
Horizon & Target & $\alpha$ & $\beta$ & $\alpha_{\sqrt{~}}$ & $\beta_{\sqrt{~}}$ & $\alpha_{\log}$ & $\beta_{\log}$ \\
\midrule
% all & all & -1.093 & 0.963 & -0.352 & 0.201 & 0.391 & 0.001\\
% \addlinespace
% all & Cases & 0.036 & 0.858 & 0.043 & 0.201 & 0.751 & -0.033\\
% all & Deaths & -0.884 & 0.868 & 0.273 & 0.121 & 0.436 & -0.023\\
% \addlinespace
% 1 & all & -1.402 & 0.923 & 0.320 & 0.088 & 0.318 & -0.014\\
% 2 & all & -1.221 & 0.967 & 0.112 & 0.164 & 0.364 & -0.003\\
% 3 & all & -1.001 & 0.984 & -0.094 & 0.241 & 0.410 & 0.008\\
% 4 & all & -0.757 & 0.986 & 0.000 & 0.299 & 0.469 & 0.015\\
1 & Cases & -0.862 & 0.876 & 0.790 & 0.087 & 0.433 & -0.024\\
2 & Cases & -0.243 & 0.877 & 0.959 & 0.162 & 0.660 & -0.031\\
3 & Cases & 0.372 & 0.855 & 1.109 & 0.238 & 0.882 & -0.037\\
4 & Cases & 0.816 & 0.837 & 1.645 & 0.296 & 1.009 & -0.036\\
\addlinespace
1 & Deaths & -1.146 & 0.832 & 0.457 & 0.048 & 0.376 & -0.035\\
2 & Deaths & -0.981 & 0.867 & 0.443 & 0.084 & 0.416 & -0.028\\
3 & Deaths & -0.807 & 0.885 & 0.349 & 0.131 & 0.453 & -0.019\\
4 & Deaths & -0.602 & 0.891 & 0.125 & 0.194 & 0.501 & -0.011\\
\bottomrule
\end{tabular}
\caption{Coefficients of three regressions for the effect of the magnitude of the median forecast on expected scores. The first regression was 
$\log[\text{WIS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], $ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
$\text{WIS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},$ where $F_{\log}$ is the predictive distribution for $\log y$. The third one was $\text{WIS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{(\text{median}(F))},$ where $F_{\sqrt{\ }}$ is the predictive distribution for $\sqrt{y}$.
}
\label{tab:HUB-regression}
\end{table}
